A Methodology and Implementation of Automated Emissions Harmonization for Use in Integrated Assessment Models

### 2017-05-24 intro.tex
results section revised

Integrated Assessment Models (IAMs) are tools used to understand the complex
interactions between energy systems, economic systems, land use, the climate,
and air pollution, among other sectors . IAMs provide global projections of
systemic change by dividing the world into a number of representative regions,
the definition of which is distinct for each model. Results from IAMs are
integral in a number of international studies, which notably include projections
of climate and economic futures. Recently, flagship IAMs have developed
scenarios based on the Shared Socioeconomic Pathways (SSPs)
\cite{van_vuuren_energy_2017, kriegler_fossil-fueled_2017, calvin_ssp4:_2017,
  fujimori_ssp3:_2017, fricko_marker_2017} which model a variety of potential
global futures. The SSPs are used in analyses underpinning the Intergovernmental
Panel on Climate Change (IPCC) and the United Nations Framework Convention on
Climate Change (UNFCCC)'s Paris Agreement \TODO{cite}.

@@

In order to incorporate the large variety of possible model results and given
relatively large increases in regional and sectoral dimensionality, an automated
process for determining harmonization methods is needed. The remainder of this
paper describes the methodology of the \code{aneris} code base, written in the
Python programming language, which implements such a system in Section
\ref{sec:meths}. The results of applying the automated harmonization mechanism
on two example IAM scenarios, one with emissions growth and another with
emissions mitigation, is presented in Section \ref{sec:results}. Finally, the
general effectiveness and potential future improvements on the automated
methodology is discussed in Section \ref{sec:future}.

>>>>

Integrated Assessment Models (IAMs) are tools used to understand the complex
interactions between energy systems, economic systems, land use, the climate,
and air pollution, among other sectors . IAMs provide global projections of
systemic change by dividing the world into a number of representative regions,
the definition of which is distinct for each model. Results from IAMs are
integral in a number of international studies, which notably include projections
of climate and economic futures. Recently, flagship IAMs have developed
scenarios based on the Shared Socioeconomic Pathways (SSPs)
\cite{van_vuuren_energy_2017, fricko_marker_2017, fujimori_ssp3:_2017,
  calvin_ssp4:_2017, kriegler_fossil-fueled_2017} which model a variety of
potential global futures. The SSPs are used in analyses underpinning the
Intergovernmental Panel on Climate Change (IPCC) and the United Nations
Framework Convention on Climate Change (UNFCCC)'s Paris Agreement \TODO{cite}.

@@

In order to incorporate the large variety of possible model results and given
relatively large increases in regional and sectoral dimensionality, an automated
process for determining harmonization methods is needed. The remainder of this
paper describes the methodology of the \code{aneris} code base \TODO{cite},
written in the Python programming language, which implements such a system in
Section \ref{sec:meths}. The results of applying the automated harmonization
mechanism on two example IAM scenarios, one with emissions growth and another
with emissions mitigation, is presented in Section \ref{sec:results}. Finally,
the general effectiveness and potential future improvements on the automated
methodology is discussed in Section \ref{sec:future}.

### 2017-05-24 results.tex
intro added
rough draft of results section
results section revised

\section{Results \& Conclusions}

\subsection{A Harmonization Test Case}

In order to test the \code{aneris} harmonization procedure, results from the IAM
MESSAGE-GLOBIOM \TODO{CITE} was processed. Two scenarios from the Shared Socioeconomic
Pathways scenario library \TODO{CITE} are investigated. The reference SSP2, or
``middle of the road'' is chosen first because MESSAGE-GLOBIOM is the marker
scenario for this SSP. Notably, emissions from multiple sectors tend to increase
in SSP2, thus testing the harmonization mechanism with a increasing emission
pathway. Additionally, the SSP-45 scenario is chosen, where the ``45''
designation refers to a scenario with end-of-century emissions resulting in
approximately a 4.5 $\frac{\text{W}}{\text{m}^2}$ warming effect by the end of
the century, due to both warming from Greenhouse Gases (GHGs) and fluoridated
gases as well as cooling from aerosols. In this scenario, mitigation
technologies and policies are enacted causing a general reduction in pollutants
and GHGs, including (eventual) negative CO2 emissions in some regions and
sectors due to carbon capture and sequestration and afforestation. This scenario
explores the harmonization mechanism's response to generally decreasing emission
trajectories.

MESSAGE-GLOBIOM includes a representation of 11 distinct regions which can be
mapped directly to the 5-region definition used in the RCPs. Historical data is
taken from CEDS and LUC \TODO{get correct name}, which comprise 10 separate
pollutant and GHG species and 13 sectors. Therefore, a total of 1430 distinct
trajectories were harmonized.

The effect of harmonization overrides is analyzed by harmonizing each scenario
first without overrides and then with overrides. Of the 1430 trajectories,
approximately 2\% required the use of harmonization overrides after an initial
investigation; thus, 98\% of all trajectories were satisfactorily determined
using the default methods as chosen by the harmonization algorithm. Figure
\ref{fig:nox} provides one such example of a emissions species and sector in
which all regions were satisfactorily harmonized with the default methods. As
can be seen, results with and without overrides are identical.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{results_NOx_Energy_Sector.pdf}
    \caption[]{
      \label{fig:nox}
      NOx Energy Sector harmonized (solid lines) and unharmonized (dashed lines)
      trajectories for SSP2 and SSP2-45 are presented. SSP-45 is denoted with
      markers. Panel \textbf{a} shows 5-Region trajectories for scenarios
      \textit{without} any overrides. Panel \textbf{b} shows 5-Region
      trajectories for scenarios \textit{with} overrides. Panels \textbf{c} and
      \textbf{d} show global trajectories without and with overrides,
      respectively.  }
  \end{center}
\end{figure}

The harmonization of emissions pathways is performed in order to accurately
represent an updated historical emissions trajectory while also maintaining
consistency with the original, unharmonized pathway. When the default methods as
provided by the harmonization procedure distort or otherwise sufficiently
misrepresent the underlying unharmonized results, an override method is required
to be provided for the trajectory of the region, sector, and species in
question. Based on the 2\% of total trajectories required to be overridden, two
classifications were observed: regional trajectories whose \textit{magnitude}
was overly distorted and regional trajectories whose \textit{shape} was overly
distorted.

Figure \ref{fig:co} presents a situation in which the magnitude of a trajectory
is distorted. A large discrepancy (~300\% relative difference) is observed in
the harmonization year for carbon monoxide (CO) emissions in the industrial
sector specifically for the South Asia (SAS) MESSAGE-GLOBIOM region, which
comprises most of the Asian subcontinent. The default method chosen
(\code{constant_ratio} maintains model trends for the region, overall model
results are distorted. By applying a \code{constant_offset} override, the
regional trend and shape is maintained. With the new harmonization method for
the SAS region, the global trajectory for industrial CO also corresponds more
closely with the unharmonized trajectory.

@@

In certain circumstances, the application of the default harmonization methods
can affect not only the magnitude but also the shape of regional
trajectories. Figure \ref{fig:nh3} shows emissions trajectories for ammonia
(NH3) resulting from the agriculture sector in Asia. Again, the SAS region shows
a large discrepancy in the harmonization year (> 150\% in this case). The
resulting trajectory harmonized with a \code{constant_ratio} method by default
provides a large increase after 2080 in the SSP2 reference scenario. Notably,
the SSP2-45 scenario is not affected to the same degree. While this distortion
affects the magnitude of the SAS trajectory, it largely affects the post-2080
shape of the global trajectory (see Figure \ref{fig:nh3}, panel \textbf{c}). By
using a \code{constant_offset} method as an override, this distortion is
addressed and more accurately reflects unharmonized results both in the SAS
region and global results for agricultural ammonia emissions.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{results_NH3_Agriculture.pdf}
    \caption[]{
      \label{fig:nh3}
      NH3 Agriculture harmonized and unharmonized emissions are presented for
      SSP2 and SSP2-45 scenarios. Scenarios and panel layouts are identical to
      Figure \ref{fig:co}. In this case, the SAS trajectory again shows not only
      a magnitude distortion, but also a shape distortion at the tail of the
      trajectory. Override methods have been applied to correct the distortion.
    }
  \end{center}
\end{figure}

\subsection{Discussion \& Future Work}

>>>>

\section{Results}\label{sec:results}

In order to test the \code{aneris} harmonization procedure, results from the IAM
MESSAGE-GLOBIOM \cite{fricko_marker_2017} was processed. Two scenarios from the
SSP scenario library \cite{Riahi2017153,Rao2017346} are investigated. The SSP2,
or ``middle of the road'', scenario family is chosen to be analyzed because
MESSAGE-GLOBIOM is the marker scenario for this SSP. Of all SSP2 scenarios, the
reference (marker) and SSP2-45 scenarios are analyzed. Notably, emissions from
multiple sectors tend to increase in the reference scenario, thus testing the
harmonization mechanism with an increasing emissions pathway. The SSP-45
scenario, where the ``45'' designation refers to a scenario with end-of-century
emissions resulting in approximately a 4.5 $\frac{\text{W}}{\text{m}^2}$
radiative forcing due to both warming from GHGs and fluoridated gases as well as
cooling from aerosols, is chosen because mitigation technologies and policies
are enacted causing a general reduction in pollutants and GHGs, including
(eventual) negative CO2 emissions in some regions and sectors due to carbon
capture and sequestration and afforestation. This scenario explores the
harmonization mechanism's response to generally decreasing emission
trajectories.

MESSAGE-GLOBIOM includes a representation of 11 distinct regions which can be
mapped directly to the 5-region definition used in the SRES. Historical data is
taken from previously described LUC and anthropogenic sources, which comprise 10
separate pollutant and GHG species and 13 sectors. Therefore, a total of 1430
distinct trajectories were harmonized for each scenario.

The effect of harmonization overrides is analyzed by harmonizing each scenario
first without and then with overrides. Of the 1430 trajectories, approximately
2\% required the use of harmonization overrides after an initial investigation;
thus, 98\% of all trajectories were satisfactorily determined using the default
methods as chosen by the harmonization algorithm. Figure \ref{fig:nox} provides
one such example of a emissions species and sector in which all regions were
satisfactorily harmonized with the default methods. As can be seen, results with
and without overrides are identical.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{results_NOx_Energy_Sector.pdf}
    \caption[]{
      \label{fig:nox}
      NOx Energy Sector harmonized (solid lines) and unharmonized (dashed lines)
      trajectories for SSP2 and SSP2-45 are presented. SSP-45 is denoted with
      ``x'' markers. Panel \textbf{a} shows 5-Region trajectories for scenarios
      \textit{without} any overrides. Panel \textbf{b} shows 5-Region
      trajectories for scenarios \textit{with} overrides. Panels \textbf{c} and
      \textbf{d} show global trajectories without and with overrides,
      respectively.  }
  \end{center}
\end{figure}

The harmonization of emissions pathways is performed in order to accurately
represent an updated historical emissions trajectory while also maintaining
consistency with the original, unharmonized pathway. As such, when the default
methods as provided by the harmonization procedure distort or otherwise
sufficiently misrepresent the underlying unharmonized results, an override
method is required to be provided for the trajectory of the region, sector, and
species in question. Based on the 2\% of total trajectories required to be
overridden, two classifications were observed: regional trajectories whose
\textit{magnitude} was overly distorted and regional trajectories whose
\textit{shape} was overly distorted.

Figure \ref{fig:co} presents a case in which the magnitude of a trajectory is
distorted. A large discrepancy ($\sim$300\% relative difference) is
observed in the harmonization year for carbon monoxide (CO) emissions in the
industrial sector specifically for the South Asia (SAS) MESSAGE-GLOBIOM region,
which comprises most of the Asian subcontinent. The default method chosen
(\code{constant_ratio}) maintains model trends for the region; however, overall
model results are distorted. By applying a \code{constant_offset} override, the
regional trend and magnitude is maintained. With the new harmonization method
for the SAS region, the global trajectory for industrial CO also corresponds
more closely with the unharmonized trajectory.

@@

In certain circumstances, the application of the default harmonization methods
can affect not only the magnitude but also the shape of regional
trajectories. Figure \ref{fig:nh3} shows an example case of emissions
trajectories for ammonia (NH3) from the agriculture sector in Asia. Again, the
SAS region shows a large discrepancy in the harmonization year ($>$150\% in
this case). The resulting trajectory harmonized with the default method
(\code{constant_ratio}) provides a large increase after 2080 in the SSP2
reference scenario. Notably, the SSP2-45 scenario is not affected to the same
degree. While this distortion changes the magnitude of the SAS trajectory, it
largely affects the post-2080 shape of the global trajectory (see Figure
\ref{fig:nh3}, panel \textbf{c}). By using a \code{constant_offset} method as an
override, this distortion is addressed and more accurately reflects unharmonized
results both in the SAS region and global results for agricultural ammonia
emissions.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{results_NH3_Agriculture.pdf}
    \caption[]{
      \label{fig:nh3}
      NH3 agricultural harmonized and unharmonized emissions are presented for
      SSP2 and SSP2-45 scenarios. Scenarios and panel layouts are identical to
      Figure \ref{fig:co}. In this case, the SAS trajectory again shows not only
      a magnitude distortion, but also a shape distortion at the tail of the
      trajectory. Additionally, global trajectories are greatly affected by the
      harmonization method choice (there is $\sim$20 percent relative difference
      between trajectories in the reference scenario in 2100). Override methods
      have been applied to correct the distortion.  }
  \end{center}
\end{figure}
